---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

##knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
setwd("C:/") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library("lubridate")
library("fpp3")
library("tsibble")
library('feasts')
library("fpp2")
library("na.tools")
library("anomalize")
```


Part 1 - Predicting/forecasting the LendingClub loans


```{r}
df = file.choose()
```



```{r}
tslCOrg <- read.csv(df)

```


```{r}
skim(tslCOrg)
```




```{r}
tslCOrg <- tslCOrg %>% 
  mutate(date=as.character(date),state = as.character(state))

```




```{r}
tslCOrg <- tslCOrg %>% 
  mutate(date = yearmonth(date)) %>% 
  as_tsibble(index = date , key = state)

```


```{r}
df1 = file.choose()
```

```{r}
dfnye <- read.csv(df1)

```

```{r}
dfnye <- dfnye %>% 
  mutate(date = as.character(date),state = as.character(state))
```




```{r}
dfnye <- dfnye %>% 
  mutate(date = yearmonth(date))

```




```{r}
df2 = file.choose()
```

```{r}
dfc <- read.csv(df2)

```

```{r}
dfstate <- data.frame(state = state.abb,NAME = state.name)

```




```{r}
dfc <- merge(x = dfc,  y = dfstate, by="NAME",all.x=TRUE)

```


```{r}
dfc <- dfc %>% 
  mutate(state = as.character(state))

```



```{r}
tslCOrg <- merge(x=tslCOrg,y=dfc, by="state",all.x=TRUE)

```

```{r}
tslCOrg %<>% 
  rowwise() %>% 
  mutate(loansPerCapita = totalLoans/CENSUS2010POP)

```


```{r}
tslc <- merge(x = tslCOrg,y = dfnye,by = c("state","date"),all.x = TRUE)

  
```




```{r}
tslc <- tslc %>% 
  as_tsibble(index = date, key = state)

```


```{r}
dfc[order(-dfc$CENSUS2010POP),]
```





```{r}
tslc %>% 
  filter(state == c("CA","TX","NY","FL","IL")) %>% 
    autoplot(loansPerCapita)
```



```{r}
tslc %>% 
  filter(state == c("SD","AK","ND","VT","WY")) %>% 
    autoplot(loansPerCapita)
```



```{r}
anoMA <- tslc %>% 
  filter(state == "MA")
anoMA <- as_tsibble(anoMA, key = state, index = date)

anoCO <- tslc %>% 
  filter(state == "CO")
anoCO <- as_tsibble(anoCO, key = state, index = date)

anoNY <- tslc %>% 
  filter(state == "NY")
anoNY <- as_tsibble(anoNY, key = state, index = date)
```

```{r}
anoMA <- anoMA %>% 
  mutate(date = as.Date(date))

```


```{r}
anoCO <- anoCO %>% 
  mutate(date = as.Date(date))

```


```{r}
anoNY <- anoNY %>% 
  mutate(date = as.Date(date))

```



```{r}
anomalyMA <- 
  anoMA %>% 
  time_decompose(loansPerCapita, method = "stl") %>% 
  anomalize(remainder, method = "iqr") %>% 
  plot_anomaly_decomposition() +
  labs(title = "Anomaly Detection for the MA data")+
  xlab("date") + ylab("Loans Per Capita")+
  scale_x_date(date_breaks = "years", date_labels = "%y")
anomalyMA
```




```{r}
anomalyCO <- 
  anoCO %>% 
  time_decompose(loansPerCapita, method = "stl") %>% 
  anomalize(remainder, method = "iqr") %>% 
  plot_anomaly_decomposition() +
  labs(title = "Anomaly Detection for the CO data")+
  xlab("date") + ylab("Loans Per Capita")+
  scale_x_date(date_breaks = "years", date_labels = "%y")
anomalyCO
```


```{r}
anomalyNY <- 
  anoNY %>% 
  time_decompose(loansPerCapita, method = "stl") %>% 
  anomalize(remainder, method = "iqr") %>% 
  plot_anomaly_decomposition() +
  labs(title = "Anomaly Detection for the NY data")+
  xlab("date") + ylab("Loans Per Capita")+
  scale_x_date(date_breaks = "years", date_labels = "%y")
anomalyNY
```



```{r}
tslc %>% 
  filter(state == "NY") %>% 
   model(STL(loansPerCapita~ season(window='periodic'),robust = TRUE)) %>% 
      components() %>% 
       autoplot()
```


```{r}
tslc %>% 
  filter(state == 'NY') %>% 
  gg_season(loansPerCapita)
```





```{r}
tslc %>% 
  filter(state == 'NY') %>% 
  gg_subseries(loansPerCapita)
```



```{r}
tslc %>% 
  filter(state=="NY") %>% 
    ACF(loansPerCapita) %>% 
          autoplot()
  
```


```{r}
tslc %>% 
  filter(state=="NY") %>% 
    PACF(loansPerCapita) %>% 
          autoplot()
```


```{r}
tslc %>% 
  filter(state=="NY") %>% 
    gg_lag(loansPerCapita, lags = c(1,5,10,15,20,25)  ,geom = "point")
```


```{r}
nyc_ma <- tslc %>% 
  filter(state == "NY") %>% 
    mutate('5-MA' = slide_dbl(loansPerCapita, mean, .size = 5, .align = "center"))
```




```{r}
nyc_ma %>% 
  autoplot(nyc_ma$loansPerCapita) +
    autolayer(nyc_ma, nyc_ma$`5-MA`, color = 'red')
```




```{r}
loan_fit <- tslc %>% 
    filter(state == 'NY') %>% 
      model(seasonalNaive = SNAIVE(loansPerCapita), DRIFT = NAIVE(loansPerCapita~drift()))

  
```

```{r}
loan_fc <-  loan_fit %>% 
  forecast(h=60)


```





```{r}
loan_fc %>% 
  autoplot(tslc %>% filter(state == "NY"))
```


```{r}
tslc_ny <- tslc %>% 
  filter(state == "NY")

```






```{r}
fit_loans <- tslc_ny %>% 
  model(TSLM(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + trend() + season()))
report(fit_loans)
```





```{r}
fitfitted <- 
  augment(fit_loans) %>%
    ggplot(aes(x=date))+
    geom_line(aes(y= loansPerCapita, color = "Data"))+
    geom_line(aes(y=.fitted, color = "Fitted"))
plot(fitfitted)
```



```{r}
fit_loans1 <- tslc_ny %>% 
  model(TSLM(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits ))
report(fit_loans1)
```

```{r}
fitfitted1 <- 
  augment(fit_loans1) %>%
    ggplot(aes(x=date))+
    geom_line(aes(y= loansPerCapita, color = "Data"))+
    geom_line(aes(y=.fitted, color = "Fitted"))
plot(fitfitted1)
```



```{r}
train1 <- tslc_ny %>% filter_index("2007 Jun" ~ "2014 Dec")
test1 <-  tslc_ny %>% filter_index("2015 Jan" ~ "2017 Mar")

```





```{r}
loan_fit <- train1 %>%
  model(TSLM(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + trend() + season()))

```

```{r}
loans_fc <- loan_fit %>% 
  forecast(test1)
loans_fc
```

```{r}
train1 %>%
ggplot() +
geom_line(aes(x = date, y = loansPerCapita, colour = "Train")) + geom_line(aes(x = date, y = loansPerCapita, colour = "Predict"), data = loans_fc) + geom_line(aes(x = date, y = loansPerCapita, colour = "Test"), data = test1) +
xlab("Date") + ylab("Loans Per Capita") +
ggtitle("Lending Club Predictive Plot 1") +
guides(colour=guide_legend(title=NULL))
```






```{r}
train2 <- tslc_ny %>% filter_index("2007 Jun" ~ "2015 Dec")
test2 <-  tslc_ny %>% filter_index("2016 Jan" ~ "2017 Mar")

```





```{r}

loan_fit1 <- train2 %>%
  model(TSLM(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + trend() + season()))

```

```{r}

loans_fc1 <- loan_fit1 %>% 
  forecast(test2)

```




```{r}
train2 %>%
ggplot() +
geom_line(aes(x = date, y = loansPerCapita, colour = "Train")) + geom_line(aes(x = date, y = loansPerCapita, colour = "Predict"), data = loans_fc1) + geom_line(aes(x = date, y = loansPerCapita, colour = "Test"), data = test2) +
xlab("Date") + ylab("Loans Per Capita") +
ggtitle("Lending Club Predictive Plot 2") +
guides(colour=guide_legend(title=NULL))
```




```{r}
fit_loans %>% 
  gg_tsresiduals()
```


```{r}
fit_arima <- tslc_ny %>% 
  model(ARIMA(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits, stepwise = FALSE, approximation = FALSE))
report(fit_arima)
  
```


```{r}
fit_arima1 <- tslc_ny %>% 
  model(ARIMA(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits))
report(fit_arima1)
```


#p-Values

```{r}
avgTerm <- 2*pt(-abs(0.0025/0.0017),nrow(tslc_ny)-18)
avgIntRate <- 2*pt(-abs(-0.0148/0.0077),nrow(tslc_ny)-18)
avgGrade <- 2*pt(-abs(0.0217/0.0167),nrow(tslc_ny)-18)
avgEmpLength <- 2*pt(-abs(0.0005/0.0064),nrow(tslc_ny)-18)
avgVerifStatus <- 2*pt(-abs(0.0131/0.0218),nrow(tslc_ny)-18)
avgHomeOwner <- 2*pt(-abs(-0.0280/0.0555),nrow(tslc_ny)-18)
NYCPI <- 2*pt(-abs(-0.0013/0.0005),nrow(tslc_ny)-18)
NYUnemployement <- 2*pt(-abs(0.00199/0.0105),nrow(tslc_ny)-18)
NYCondoPriceIdx <- 2*pt(-abs(0.003/0.001),nrow(tslc_ny)-18)
```

```{r}
avgTerm
avgIntRate
avgGrade
avgEmpLength
avgVerifStatus
avgHomeOwner
NYCPI
NYUnemployement
NYCondoPriceIdx
```




```{r}
tslc_ny %>% 
  features(loansPerCapita, unitroot_ndiffs)
```




```{r}
fit_arima1 <- tslc_ny %>% 
  model(ARIMA(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits+ pdq(2:6,1,0:3) ,stepwise = FALSE, approximation = FALSE))
report(fit_arima1)
  
```



```{r}
set.seed(333)
dfTrain1 <- tslc_ny %>% filter_index("2007 Jun" ~ "2016 Feb")
dfTest1 <- tslc_ny %>% filter_index("2016 Mar"~"2017 Mar")
  
```





```{r}
test_fit <- dfTrain1 %>% 
  model(TS1 = TSLM(loansPerCapita ~ trend() + season() ),TS2 = TSLM(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + trend() + season()), ARIMA1 = ARIMA(loansPerCapita), ARIMA2 = ARIMA(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + pdq(4,0,0) + PDQ(1,0,0) ) )


```

```{r}
glance(test_fit)
```


```{r}
test_fc <- test_fit %>% 
  forecast(dfTest1)
```





```{r}
accuracy(test_fc, dfTest1)
```




```{r}
set.seed(333)
dfTrain2 <- tslc_ny %>% filter_index("2007 Jun" ~ "2016 Mar")
dfTest2<- tslc_ny %>% filter_index("2016 Apr"~"2017 Mar")
```




```{r}
test_fit1 <- dfTrain2 %>% 
  model(TS1 = TSLM(loansPerCapita ~ trend() + season() ),TS2 = TSLM(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + trend() + season()), ARIMA1 = ARIMA(loansPerCapita), ARIMA2 = ARIMA(loansPerCapita ~ avgTerm+ avgIntRate+ avgGrade+ avgEmpLength+ avgAnnualInc+ avgVerifStatus+ avgHomeOwner + avgRevolBal + countOfLoans + NYCPI + NYUnemployment + NYCondoPriceIdx + NYSnapBenefits + pdq(4,0,0) + PDQ(1,0,0) ) )
```

```{r}
test_fc1 <- test_fit1 %>% 
  forecast(dfTest2)
```





```{r}
accuracy(test_fc1, dfTest2)
```




















